An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications

### 2014-05-07 paper.tex
more changes

At a more detailed view, the various layers of the IO stack each contribute
different functionality and performance implications.  The
architecture (Figure~\ref{fig:arch-mapping}) incorporates five layers, some of
which have potentially optional components.  The top layer is generally a high
level IO library, such as the demonstration HDF-5 library~\cite{hdf5}. It is in
dark blue. The intent is to only have access to the storage stack through such
an API to manage the complexity of working with the lower layers and to enable
advanced functionality that would require more direct user intervention. This
layer incorporate the necessary features for ACGs from a end-user's
perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack and is the only way to reach the persistent
storage array in lower layers (in green). Ideally, the IOD layer's
functionality can be optional based on available hardware and provided on the
IO Nodes (IONs). Much of the functionality offered at this layer would shift
either up or down the stack as discussed in detail below. The Distributed
Asynchronous Object Storage (DAOS) layer serves as the persistent storage
interface and translation layer between the user-visible object model and the
requirements of the underlying storage infrastructure. It is intended to be the
traditional file system-like foundation on which everything else is built with
no dependence on any technologies specified above it (in pink and yellow). For
example, the IOD layer with or without burst buffers is not required for DAOS
to operate properly.  Instead, the DAOS layer can handle all of the IO
operations from the user API layer, albiet with the potential performance
penalty of manipulating the shared, persistent storage array. At the bottom is
the Versioning Object Storage Device (VOSD) (in purple).  It serves as the
interface for storing objects of all types efficiently for each storage device
in the parallel storage array. Think of this layer as the physical disk
interface layer. In terms of Lustre, this would replace the API on individual
storage devices with an interface friendlier to the containers of objects and
transactions/epochs concepts used in the higher layers.

@@

The major previous work is really limited to full file systems rather than the
mountain of file system refinements made over the years.

@@

PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle.

Lustre~\cite{braam:lustre-arch} has become the defacto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system. Scalability
limitations are part of the motivation for the FFSIO project.

@@

The idea of sharding is to store a portion of a data set on a single storage
target. This is similiar in concept to the HDF-5 hyperslab. The FFSIO stack
supports sharding the data in the default or some other structured way as well
as ``re-sharding'' based on application needs. For example, reordering the
data so that a different dimension is the ``fast'' dimension may greatly
improve the performance of a subsequent data analytics task. A common scenario
where this is common is a Fortran code (column-major) writes data for a C code
(row-major) to analyze.

@@

\section{VOSD Layer}\label{sec:vosd}
The Versioning Object Storage Device (OSD) layer operates as the layer for each
persistent storage device used to support the parallel storage array. In the
purest form, it uses a local file system to arrange storage of objects that
represent parts of the higher level objects in containers.

@@

@@

This will exemplify how applications make use of the Exascale stack. For FF,
support for GraphLab and GraphBuilder will be prototyped, which are
graph-processing frameworks. GraphBuilder is a set of MapReduce tasks that
extract, normalize, partition and serialize a graph out of unstructured data,
and writes graph-specific formats into HDFS. These files are later consumed by
GraphLab, a vertex-centric, asynchronous execution engine that runs directly on
top of HDFS (i.e., non-MapReduce). The following illustrates the architecture
of both frameworks:

@@

In order to make both work on top of the exascale stack, both have to be
modified. After these modifications are implemented, GraphBuilder will
be able to write the partitioned graph in (the newly proposed) HDF5
files which will thus be stored in the IOD nodes (or IONs) in a
parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
the library to perform at high speeds by benefiting from the new
features (see next section). In general both frameworks will be modified
so that calls to HDFS-based formats are replaced by the proposed HDF5
ones. This is referred to as the HDF Adaptation Layer or HAL and will
provide, from the GraphBuilder/GraphLab point of view:

@@

\section{Demonstration}
This stack has a prototype implementation intended to test concepts rather
than performance and scalability. It has focused on examining the interaction
of the different APIs for each layer to flesh out any detailed requirements
or concerns that may have been missed in the conceptualization of this IO stack.

@@

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism. The
split between the IOD layer and the DAOS layer offers a fast place for
intermediate data without requiring the overhead of writing to persistent
storage. The envisioned transaction mechanism, while not perfect in the current
form, is another good attempt to address both failures and prevent access to
incomplete or incorrect data by downstream data consumers. Integrated with the
IOD functionality, this concept represents the consensus approach for what will
be required.

The partial metadata management incorporated into the IOD layer and the lack of
consideration for how to handle and recover from failures are oversights in the
current documents. It is our understanding that these will be addressed in the
next phase and we hope to help inform that effort with our experiences.

We hope that the efforts made in the \DDTns, Lightweight File Systems, and
other efforts to explore the requirements for this space, along with the
analysis presented in this paper will prove useful for the next phase of the
Fast Forward project.

>>>>

At a more detailed view, the various layers of the IO stack each contribute
different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
incorporates five layers, some of which have potentially optional components.
The top layer is generally a high level IO library, such as the demonstration
HDF-5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers and to enable advanced functionality that would
require more direct user intervention. This layer incorporates the necessary
features for ACGs from a end-user's perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack (in green) and offers features like Burst
Buffers to insulate the persistent storage array from bursty IO workloads.
Ideally, the IOD layer's functionality can be optional based on available
hardware and compute power provided on the IO Nodes (IONs). Much of the
functionality offered at this layer would shift either up or down the stack as
discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
layer serves as the persistent storage interface and translation layer between
the user-visible object model and the requirements of the underlying storage
infrastructure. It is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albiet with the potential performance penalty of manipulating
the shared, persistent storage array. At the bottom is the Versioning Object
Storage Device (VOSD) (in purple).  It serves as the interface for storing
objects of all types efficiently for each storage device in the parallel
storage array. Think of this layer as the physical disk interface layer. In
terms of Lustre, this would replace the API on individual storage devices with
an interface friendlier to the containers of objects and transactions/epochs
concepts used in the higher layers.

@@

The major previous work is really limited to full file systems rather than the
mountain of file system refinements made over the years. A selection of these
other file systems and some features that make it relatively unique are
described below.

@@

PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle. It has been
commercialized in recent years as OrangeFS.

Lustre~\cite{braam:lustre-arch} has become the defacto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.

@@

The idea of sharding is to store a logically complete portion of a data set on
a single storage target. This is similiar in concept to the HDF-5 hyperslab.
The FFSIO stack supports sharding the data in the default or some other
structured way as well as ``re-sharding'' based on application needs. For
example, reordering the data so that a different dimension is the ``fast''
dimension may greatly improve the performance of a subsequent data analytics
task. A common scenario where this is common is a Fortran code (column-major)
writes data for a C code (row-major) to analyze.

Additionally, sharding at the IOD layer is done in a couple of different ways.
When writing from the compute nodes, the default layout will reflect what the
processes write. This can be reorganized using function shipping. At the DAOS
layer, this data organization may change based on the type of storage devices
beneath it.

@@

\section{VOSD Layer}
\label{sec:vosd}

The Versioning Object Storage Device (OSD) layer operates as the layer for each
persistent storage device used to support the parallel storage array. In the
purest form, it uses a local file system to arrange storage of objects that
represent parts of the higher level objects in containers.

@@

Several concepts crosscut many of these layers and are best described in a
single location. For example, transactions and epochs are visible from the
user API level down into the DAOS layer. While each layer affects the concept,
it is best to look at it across all of the layers.

In the subsections that follow, we examine transactions and epochs, metadata
management, function and analysis shipping, and the aribtrary connected graphs
support.

@@

What people popularly consider Big Data applications fall into two broad
categories. First, data processing tasks that fit into the MapReduce model
where data is tagged and sorted to discover relationships. These sorts of
applications only require scale out rather than scale up. Scaling out requires
replicas to process data simultaneously, but do not need to coordinate for that
data processing. Scaling up, what scientific simulations do, requires sometimes
serious coordination between the data processing processes for any of them to
succeed. The second category is more similar to scientific simulations in that
it requires information from other processes to succedd. These graph-style
applications can be done using Hadoop, but it does not fit as well.

To address these graph applications, GraphLab and GraphBuilder are evaluated
to determine how to extend the IO stack to better support these applications.
GraphBuilder is a set of MapReduce tasks that extract, normalize, partition and
serialize a graph out of unstructured data, and writes graph-specific formats
into HDFS. These files are later consumed by GraphLab, a vertex-centric,
asynchronous execution engine that runs directly on top of HDFS (i.e.,
non-MapReduce). The following illustrates the architecture of both frameworks:

@@

In order to make both work on top of the exascale stack, both have to be
modified. After these modifications are implemented, GraphBuilder will
be able to write the partitioned graph in the newly proposed HDF5
files which will thus be stored in the IOD nodes (or IONs) in a
parallel-optimized way. On the GraphLab side, HDF5-awareness will allow
the library to perform at high speeds by benefiting from the new
features, such as the function shipping. In general both frameworks will be
modified so that calls to HDFS-based formats are replaced by the proposed HDF5
ones. This is referred to as the HDF Adaptation Layer or HAL and will
provide, from the GraphBuilder/GraphLab point of view:

@@

The initial phase of this project has determined the necessary changes in the
HDF-5 format to support these features. These identified features will be
proven during phase 2 with a demonstration of GraphBuilder and GraphLab.

\section{Demonstration}
\label{sec:evaluation}

This stack has a prototype implementation intended to test concepts rather
than performance and scalability. It has focused on examining the interaction
of the different APIs for each layer to flesh out any detailed requirements
or concerns that may have been missed in the conceptualization of this IO stack.

@@

The Fast Forward Storage and IO Stack project has designed a good first pass at
addressing the requirements for an extreme scale data storage mechanism.  By
prefering a high level user API like HDF5 rather than using the POSIX
interface, more advanced functionality can be incorporated with less end-user
impact. The introduction of the IOD layer with buffering will absorb the
difference between the compute node IO demands and the available bandwidth in
the storage array. With DAOS supporting translating the container and object
model to the underlying storage options, different storage technologies can be
deployed over time.

With the overall stack design complete with a prototype implementation,
refinements, such as fault detection and recovery, can be designed and tested.
These and other activities for phase 2 will ultimately generate what is likely
to be the next generation storage stack for extreme scale platforms.
